---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r read, message = F, warning = F}
activity <- read.csv('activity.csv')
library(ggplot2)
library(dplyr)
library(knitr)
```

## What is mean total number of steps taken per day?

```{r mean_totes, results='asis'}
hist_act <- activity %>% group_by(date) %>% summarize(Daily_Steps = sum(steps, na.rm = TRUE))
hist_act$date <- as.Date(hist_act$date, format = '%Y-%m-%d')
g <- ggplot(hist_act, aes(x=date, y = Daily_Steps)) +geom_bar(stat = 'identity')

summary_act <- activity %>% group_by(date) %>% summarize(mean_steps = mean(steps, na.rm = TRUE), median_steps = median(steps, na.rm = TRUE))
```

1. Calculate the total number of steps taken per day
```{r echo = FALSE, results='asis'}
kable(hist_act)
```

2. Histogram of the total number of steps taken each day
```{r echo = FALSE, results='asis'}
g
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r echo = FALSE, results='asis'}
kable(summary_act)
```



## What is the average daily activity pattern?
```{r avg_daily, results='asis'}
int_act<-activity %>% group_by(interval) %>% summarize(Avg_Daily_Pattern = mean(steps, na.rm = TRUE))
daily <- ggplot(int_act, aes(x=interval, y = Avg_Daily_Pattern)) + geom_line()
```

1. Time series plot of the 5-minute interval and avg number of steps taken across all days
```{r echo=FALSE,results='asis'}
daily
```

2. 5-min interval, on average across all days in the data, contains the maximum number of steps?\hfill\break
```{r echo=FALSE,results='asis'}

int_act[which(int_act$Avg_Daily_Pattern == max(int_act$Avg_Daily_Pattern)),]$interval
```


## Imputing missing values
```{r impute, results='asis'}
total_missing <- sum(is.na(activity))
impute_activity <- activity
for (i in 1:nrow(impute_activity)){
  
  if (is.na(activity[i,]$steps)){
    
    impute_activity[i,]$steps <- int_act[which(int_act$interval == impute_activity[i,]$interval),]$Avg_Daily_Pattern
    
  }
  
}

impute_activity_hist <- impute_activity %>% group_by(date) %>% summarize(Daily_Steps = sum(steps, na.rm = TRUE))
impute_activity_hist$date <- as.Date(impute_activity_hist$date, format = '%Y-%m-%d')
imp_g<-ggplot(impute_activity_hist, aes(x=date, y = Daily_Steps)) +geom_bar(stat = 'identity')

summary_impute_act <- impute_activity %>% group_by(date) %>% summarize(mean_steps = mean(steps, na.rm = TRUE), median_steps = median(steps, na.rm = TRUE))

```

1. Calculate the total number of missing values in the data set\hfill\break
```{r echo=FALSE,results='asis'}
total_missing
```

2/3. Impute missing values in the data set. Create a new data set.\hfill\break
```{r results='asis'}
impute_activity <- activity
for (i in 1:nrow(impute_activity)){
  
  if (is.na(activity[i,]$steps)){
    
    impute_activity[i,]$steps <- int_act[which(int_act$interval == impute_activity[i,]$interval),]$Avg_Daily_Pattern
    
  }
  
}
```


4. Histogram of the total number of steps taken each day: I have found the values to differ from the first part of the assignment. The data is less skewed in my opinion.\hfill\break


```{r echo=FALSE,results='asis'}
imp_g

kable(summary_impute_act)
```




## Are there differences in activity patterns between weekdays and weekends?
```{r diff, results='asis'}

impute_activity$day_type <- ifelse(weekdays(as.Date(impute_activity$date, format = '%Y-%m-%d')) == 'Saturday' | 
                                     weekdays(as.Date(impute_activity$date, format = '%Y-%m-%d')) == 'Sunday', 'weekend', 'weekday')


int_impute_act <- impute_activity %>% group_by(interval, day_type) %>% summarize(Avg_Daily_Pattern = mean(steps, na.rm = TRUE))


wk_act <- ggplot(int_impute_act, aes(x=interval, y = Avg_Daily_Pattern)) + geom_line() + facet_grid(as.factor(int_impute_act$day_type))
```


1. New factor variable with two levels for weekday and weekend
```{r results='asis'}

impute_activity$day_type <- ifelse(weekdays(as.Date(impute_activity$date, format = '%Y-%m-%d')) == 'Saturday' | 
                                     weekdays(as.Date(impute_activity$date, format = '%Y-%m-%d')) == 'Sunday', 'weekend', 'weekday')

```

2. Panel plot of the 5 minute interval (x-axis)
```{r echo=FALSE,results='asis'}
wk_act
```

